Capon Beamforming Applied on Moving Objects - An Analysis of Local Lateral Shift-Invariance

### 2013-12-10 HRUIofMovingObjects/jrnl_article.tex
Moving objects: done some writing

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The adaptive nature of the method has been shown to greatly improve the lateral resolution in ultrasound images, but with a high computational cost. In recent work we introduced an implementation of the Capon beamformer capable of processing a stream of images at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

@@

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge definitions in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if e.g. the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying a steering vector uncertainty set \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. A steering vector uncertainty sets helps to reduce signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In this paper we will make use of a set of steering vectors to reduce this self-nulling effect by means of over sampling.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to two times the two-way system resolution ($1/64$ radians) using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the scanned output when the object is located at plus and minus half the system resolution. No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to a signal-to-noise ratio (SNR) of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with a beamspacing equal to the system resolution, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $27$ dB. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam density is increased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the system resolution, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.}
\label{fig:das_capon_beams}
\end{figure}

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criterion. The minimization problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The lateral sampling required in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the combined aperture size on transmit ($D_{tx}$) and receive ($D_{rx}$), and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound in the body. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{D_{tx} + D_{rx}}\frac{1}{p}, \label{eq:resolution}
\end{align}
where the over sampling factor $p$ can be used to reduce folding of frequencies between $f_c$ and $f_c + B/2$. A $p$ factor of one has been used throughout this paper.

@@

A given imaging system will then be locally laterally shift-invariant if the contours are diagonal and if the amplitude deviation among the maximum points is less than 1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will hopefully not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse with a $\text{f-number}=3$ and corresponding cosine apodization on receive. The transmit focus in azimuth and elevation was configured to be equal and the point scatterer was moved through this focus. The lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located exactly on this beam at one time instance.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the geometric distortion and scalloping loss is increased to a level where the contours no longer are diagonal and the maximum trace is not constant. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. As mentioned, the beam at zero degrees is the only one where the point scatterer is located in the center of the beam. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal when the point scatterer has zero lateral offset. From the plot we observe that the signal cancellation is high for all the other offsets. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

@@

\section{Oversampling Methods}\label{sec:methods}
We will now discuss different ways of improving the local lateral shift-invariance when the Capon beamformer is used. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the system frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
The straight forward way of obtaining sufficient sampling in an active system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. cardiac ultrasound imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

@@

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by using phase rotation. This is also possible in broadband applications as long as the phase rotation is limited to a fraction of the wave length (See coarse-fine beamforming in \cite{Thomenius} \comment{find a better citation}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing to large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For a uniform linear array located along the x-axis the steering vector $\vec{a}_\theta$ should be calculated as 
\begin{align}
\vec{a}_\theta = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swiped from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element array centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch, $x_i - x_{i-1}$, as the full array. The steering error introduced by this steering vector for a frequency band $B$ is less than $B/2 \times \Delta/2$.

@@

Since the phase rotation method operates on the same data as before, there will be no reduction in the acquisition frame rate by applying this method. In addition, since the same number of covariance matrices are constructed and inverted, there is no extra cost involved here as well. The only additional cost will be the multiplication of the extra steering vectors with the inverted covariance matrix. If the equation $\mat{\hat{R}}\vec{a} = \vec{b}$, as in \cite{Asen}, is solved using the Gauss-Jordan algorithm, we can solve the problem for all steering vectors at once by row reducing the matrix $[\mat{\hat{R}},  \vec{a}_{\theta_1}, \vec{a}_{\theta_2}, \cdots \vec{a}_{\theta_{N_\theta}}]$ to the following form $[\mat{I},  \vec{b}_{\theta_1}, \vec{b}_{\theta_2}, \cdots \vec{b}_{\theta_{N_\theta}}]$. This has lower complexity than inverting $\mat{\hat{R}}$, $O(L^3)$, as long as $N_\theta < L$.

@@

In Fig\,\ref{fig:image_das} one point scatterer has been selected (white box), and then used to generate LSV-plots. \comment{Todo: Add lateral coherent interpolation as done for the simulations? As for now speckle is not fully resolved without it.} Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is therefore locally laterally shift-invariant. As a consequence point scatterers has a consistent appearance when subjected to motion \multimedia{Video DAS}.

Figure \ref{fig:phantom_capon_LSV} shows the same for the Capon beamformer without oversampling. In the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we get a high amplitude value at this location. In almost all the other frames the signal loss is massive. The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

Figure \ref{fig:phantom_capon_LSV_10PR} presents the result of applying ten times oversampling by phase rotation. We observe how the amplitude of the point scatterer is better preserved in all frames, and the scalloping loss is not far from our 1 dB threshold. Note that the oscillation along the diagonal does not indicate shift-variance. It is a result of lack of control when the probe was manually dragged across the phantom. In order to get a continues straight diagonal the probe has to be dragged across the phantom with a speed which match the acquisition frame rate. In addition, due to the increase in lateral resolution the Capon beamformer requires smaller frame-to-frame movements than DAS. Looking at the video of Capon beamforming with ten times oversampling by phase rotation \multimedia{Capon + 10PR} there is no visible blinking anymore. We therefore conclude that the system is now locally laterally shift invariant.

@@

\section{Discussion}\label{sec:dis}
In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root of cause was found to be a too low lateral sampling density. In Section \ref{sec:methods} we therefore presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity, was based on phase rotation. In Section \ref{sec:res} we presented results showing that this method also works on \textit{in-vitro} data when the probe is subjected to motion.

Note that in this paper we have investigated a single array configuration and a limited parameter range for the Capon beamformer. The goal was not to make a reference for which oversampling factor to use in different scenarios, but rather do a detailed investigation of the scalloping loss effect. We hope with this investigation that future readers will be able to find out them self what oversampling factor to use for their system.

If the scene contains bright point scatterers or sharp edges we recommend to measure the ratio between highest peak and the neighbouring background level. In his seminal paper \cite{Cox1973} Cox derives formulas for how large the oversampling has to be for a given output signal-to-noise ratio (SNR). E.g 47 dB is suppose to result in a ten time improvement in resolution. Be aware that Cox measured resolution as the saddle point between two impinging plane waves, which typically shows smaller scalloping loss than a single moving point scatterer. In our example more than ten times oversampling was found to be required for a 40 dB output SNR in simulations, however \textit{in-vitro} ten times oversampling was enough for the system to be locally laterally shift-invariant. Possibly due to a higher initial noise level before diagonal loading was applied.

Relating the presented phase rotation method to earlier work we find that it is similar to a SVUS. However, where SVUSs typically has a criteria for selecting the best output after applying all the steering vectors contained in the set, oversampling with phase rotation outputs all. It can also be though of as a narrow band approach to PRB where Capon is applied on all receive beams.

Oversampling by phase rotation can therefore introduce similar block artifacts as PRB, but on a smaller scale when PRB is used to lower the number of tx-beams. These artifacts could be removed in the same way as for PRB if we distribute our steering angles as beams are spread out for the STB technique. We then get a phase rotation version of STB where e.g. eight phase rotated beams are averaged into four. The geometric distortion, or stripes, seen at the bottom of the sector in the video \multimedia{Capon + 10PR} should be removed by this technique.

>>>>

\IEEEPARstart{I}{n} the last decade several authors have applied the Capon beamformer (also known as the minimum variance beamformer) for medical ultrasound imaging \cite{Synnevag2007, Vignon2008, Viola}. The adaptive nature of the method has been shown to greatly improve the lateral resolution and edge sharpness in ultrasound images, but with a high computational cost. In recent work we introduced an implementation of the Capon beamformer capable of processing a stream of images at interactive frame rates \cite{Asen}. It is therefore an interesting question how a moving object appears after Capon beamforming has been applied. Earlier work on Capon beamforming for medical ultrasound imaging has investigated the method's capabilities in single frames only, and temporal behaviour and effects has therefore been neglected.

@@

As mentioned, Capon beamforming has been shown to provide improved lateral resolution and edge sharpness in ultrasound images \cite{Synnevag2007, Synnevag2009, Chen2011}. The method accomplishes this by adjusting the array weights based on element-to-element covariance. Hence, the method yields data dependent weights. Capon beamforming is however also known for its suboptimal performance if e.g. the sample size available for covariance estimation is limited \cite{Mestre2006}, if correlated signals are present \cite{Widrow1982}, and if the assumed steering vector does not match the signal propagation vector \cite{Wax1996, Wax1996a}. This has led to the development of several robust versions of the Capon beamformer with e.g. spatial smoothing \cite{Shan1985} or diagonal loading \cite{JianLi2003} of the covariance estimate, or by applying a steering vector uncertainty set \cite{Lorenz2005, Rubsamen2013}. Spatial smoothing and diagonal loading, together with temporal smoothing of the covariance estimate, have been applied in earlier work on Capon beamforming for medical ultrasound imaging \cite{Synnevag2009}. The first two trade resolution for robustness and temporal smoothing helps to preserve speckle statistics for high resolution parameters \cite{Synnevag2007a}. However, the issue of having a mismatch between the assumed steering vector and the signal propagation vector has not been addressed directly.

A steering vector uncertainty set helps to reduce signal-nulling which appears if the signal of interest is present during the covariance estimation process and if this estimate is later used with a steering vector that does not match the signal propagation vector exactly. This self-nulling is actually how the Capon beamformer obtains its super resolution capabilities, but the signal can also get lost in the process. In \cite{JianLi2003} it was shown that most methods based on steering vector uncertainty sets are equivalent to diagonal loading where the loading factor depends on the steering vector uncertainty. In that way they make up for a given uncertainty by making the Capon beam wider. Hence, the maximal achievable resolution is decreased if the initial steering vector spacing was too coarse. In medical ultrasound imaging this initial set of steering vectors is govern by the number of receive beams. In this paper we will make use of an additional set of steering vectors to reduce the self-nulling effect by means of over sampling. The maximal achievable resolution is then maintained without the presence of severe self-nulling.

In \cite{Cox1973}, Cox derives formulas for the resolution of the Capon beamformer, and presents in great detail how the amount of mismatch between the selected steering vector and the signal propagation vector impacts the beamformer output in simple situations. Cox states the following important fact about the Capon beamformer (the $\vec{k}_3$ processor): 
\begin{quote}
"\textit{... a $\vec{k}_3$ processor requires more closely spaced beams than a ... $\vec{k}_1$-processor in order to avoid serious signal suppression effects being introduced on signals arriving from directions between the beams}", 
\end{quote}
where $\vec{k}_1$ is the delay-and-sum (DAS) beamformer. In Fig.\,\ref{fig:das_capon_beams} we have plotted the two-way response for DAS and the Capon beamformer using (26) and (32) in \cite{Cox1973}. A 64-element linear array with $\lambda/2$ element spacing, where $\lambda$ is the wavelength, is scanned over a sector equal to two times the required beam spacing ($1/64$ radians) using narrow band excitation. Located in the farfield of the sector is an object moving parallel to the array surface. The plot shows the scanned output when the object is located at plus and minus half the required beam spacing for DAS (\ref{eq:resolution}). No spatially white noise is present. Hence, the interference-noise matrix ($\sigma_0^2\mat{Q}$ in \cite{Cox1973}) is the identity. The effect of typical subarray averaging and diagonal loading ($L=32$ and $d=1/100$) has also been added to the Capon beamformer output. A diagonal loading of $1/100$ corresponds to a signal-to-noise ratio (SNR) of 20 dB before array gain is applied. The figure shows that a DAS beamformer, with the required beam spacing, could yield a $1.9$ dB scalloping loss \cite{Harris1978} if the object is moving parallel to the array. However if the Capon beamformer is used with the same beam density the scalloping loss could be as large as $27$ dB. A $1.9$ dB scalloping loss for the Capon beamforming is obtained for this specific example when the beam density is increased with a factor of 25.

\begin{figure}[!t]
\centerline{
\includegraphics[width=\linewidth]{\img fig1_2way.eps}
}
\caption{Theoretical drop in output power mid-way between receive beams for delay-and-sum and Capon beamforming ($M=64$, $L=32$, $d=0.01$) applied in a farfield-narrowband setting. %The scanned output from both beamformers is plotted for two different time instances. For the first instances an object is located at an angle equal to half the required beam spacing for delay-and-sum, and for the latter the object is moved an angular distance corresponding to the system resolution. If the beam density was reduced to one beam per system resolution the maximum scalloping loss is found at zero degrees.
}
\label{fig:das_capon_beams}
\end{figure}

@@

\subsection{Capon Beamforming}
The Capon beamformer produce a weight vector, like the one in (\ref{eq:das}), based on the impinging signal and an optimization criterion. The problem is as follow \cite{Capon1969}:
\begin{align}
&\min_{\vec{w}} E\{|z[n]|^2\} \rightarrow \min_{\vec{w}} \vec{w}^H \mat{\hat{R}} \vec{w} \label{eq:capon_optimization_criteria} \\
&\text{subjected to } \vec{w}^H\vec{a} = 1,
\end{align}
where $\mat{\hat{R}}$ is a sample covariance matrix. Hence, the resulting weight vector will minimizes the output power while maintaining unit gain in the steering direction $\vec{a}$. %The covariance matrix is usually unknown and has to be estimated from the element data. The covariance matrix in (\ref{eq:capon_optimization_criteria}) is therefore substituted with the sample covariance matrix.

@@

The lateral sampling required in an ultrasound image for a given center frequency, $f_c$, and bandwidth, $B$, is given by the aperture size, $D$, and the wavelength of the center frequency, $\lambda_c = c/f_c$, where $c$ is the speed of sound in the body. The required Nyquist beam spacing is then \cite{Hergum2007}:
\begin{align}
\Delta < \frac{\lambda_c}{pD}, \label{eq:resolution}
\end{align}
where the over sampling factor $p$ can be used to reduce folding of frequencies between $f_c$ and $f_c + B/2$. A $p$-factor of two has been used throughout this paper.

@@

A given imaging system will then be locally laterally shift-invariant if the contours are diagonal and if the amplitude variation among the maximum points is less than 1 dB. Then there is minimal geometric distortion (low P-MAE) and the scalloping loss will hopefully not be visible for the observer (low A-MAE). %Since both effects can occur at the same time it is important to have a measure not effected by geometric distortion.

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). The dotted vertical lines indicates the position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} with a probe setup equal to the M5S probe by GE Vingmed Ultrasound, Horten, Norway. The probe was excited with a 2.5 MHz and 1.5 cycles long pulse with a $\text{f-number}=3$ and corresponding cosine apodization on receive. The transmit focus in azimuth and elevation was configured to be equal and the point scatterer was moved through this focus. The lateral sampling density has been doubled prior to detection through coherent interpolation. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

@@

In Fig.\,\ref{fig:das_shiftvariant} the beam density is reduced by a factor of two. We see how the geometric distortion and scalloping loss is increased to a level where the contours no longer are diagonal and the maximum trace has larger deviations. Hence, the imaging system is no longer locally laterally shift-invariant. We observe that the maximum peak is both geometrically distorted and attenuated when the point is located between two beams and the distance between the beams are too high.

@@

\section{Local Lateral Shift-Invariance of the Capon Beamformer}\label{sec:capon_LLSI}
We will now investigate the local lateral shift-invariance of the Capon beamformer with parameters equal to the ones used in earlier work \cite{Synnevag2009, Asen}. In Fig.\,\ref{fig:capon1} the same point scatterer response as presented in Fig.\,\ref{fig:das_shiftinvariant} has been weighted with Capon weights calculated using (\ref{eq:capon_weights}) with $L = M/2 = 48$, $K=1$ and $d=1/100$. As mentioned, the beam at zero degrees is the only one where the point scatterer is located in the center of the beam. The signal cancellation caused by the difference between the signal propagation vector and the assumed steering vector is therefore minimal when the point scatterer has zero lateral offset. From this plot we observe that the signal cancellation is high for all the other offsets. As a consequence the scalloping loss is tremendous, and not far from the predicted value in Fig\,\ref{fig:das_capon_beams}. %The difference between the predicted and the simulated loss might be caused by several factors. The most important one being that Fig.\,\ref{fig:das_capon_beams} shows a single frequency whereas the RMS operator used in Fig.\,\ref{fig:capon1} average a broadband signal in range. For a slice through the center of the pulse the scalloping loss is below -30dB.

@@

\section{Oversampling Methods}\label{sec:methods}
We will now discuss different ways of improving the local lateral shift-invariance when the Capon beamformer is applied. From the previous section it should be clear that we somehow need to increase the lateral sampling to make the Capon imaging system locally laterally shift-invariant. In this section we will investigate how this oversampling should be conducted, how large it needs to be and how the proposed methods will affect the system frame rate and the processing complexity.

\subsection{Oversampling on Transmit}
A straight forward way of obtaining sufficient sampling in an active imaging system is by increasing the number of transmit beams (tx-beams) until the system is locally laterally shift-invariant. In Fig.\,\ref{fig:capon_oversampling_a} and Fig.\,\ref{fig:capon_oversampling_b} we have increased the lateral sampling on transmit by a factor of four and ten respectively. As expected the geometric distortion and scalloping loss decreases with increased lateral sampling. Note that Fig.\,\ref{fig:capon_oversampling_b} is zoomed compared to Fig.\,\ref{fig:capon} in order to better see the beam-to-beam variation.

@@

\subsection{Oversampling by Parallel Receive Beamforming}
Parallel receive beamforming (PRB) means that multiple narrow receive beams (rx-beams) are constructed from one broad tx-beam. Then, by reducing the number of tx-beams and relying on multiple rx-beams in order to get sufficient sampling on receive, it is possible to achieve high frame rates in e.g. 3D cardiac ultrasound imaging. However, PRB also introduces geometrical distortion which to some degree can be compensated for by using e.g. synthetic transmit beamforming (STB) \cite{Hergum2007, Denarie2011, Denarie2013}. In \cite{Rabinovich2013} it is shown that STB works with Capon beamforming. In \cite{Asen} we also used STB in order to increase the beam density of a cardiac ultrasound recording. In the following discussion we consider corrected PRB as being equal to oversampling on transmit when it comes to image quality. The LSV-plots for PRB are therefore not included.

@@

\subsection{Oversampling by Phase Rotation}
The Capon beamformer formula in (\ref{eq:capon_weights}) gives a direct opportunity for oversampling using phase rotation. The steering vector $\vec{a}$ has in previous work been set to $\vec{1}$ since the ultrasound data are pre-delayed. However, the steering vector can also, as in narrow band applications, be varied over a set of pre-defined vectors. In a narrow band setting the beam can make a sweep across the whole sector by using phase rotation. This is also possible in broadband applications as long as the phase rotation is less than one wave length (See coarse-fine beamforming in \cite{Thomenius}). It is therefore possible to make a sweep between a given beam and its two nearest neighbours using only phase rotation without introducing large errors. Phase rotation is actually what the Capon beamformer use internally to micro-steer away from bright point scatterers and to enhance edges.

For a uniform linear array located along the x-axis the steering vector $\vec{a}_\theta$ should be calculated as 
\begin{align}
\vec{a}_\theta = 
\begin{bmatrix}
e^{-j\frac{2\pi}{\lambda_c}x_0\sin(\theta)} \\
e^{-j\frac{2\pi}{\lambda_c}x_1\sin(\theta)} \\
\vdots \\ 
e^{-j\frac{2\pi}{\lambda_c}x_{L-1}\sin(\theta)}
\end{bmatrix}
\end{align}
where $x_i$ is the element position and $\theta$ is swept from $-\Delta/2$ to $\Delta/2$ where $\Delta$ is as defined in (\ref{eq:resolution}). Note that the steering vector is calculated for a $L$ element subarray centered around the origin, even though the full array has $M$ elements. It is important that this subarray has the same element pitch, $x_i - x_{i-1}$, as the full array. %The steering error introduced by this steering vector for a frequency band $B$ is less than $B/2 \times \Delta/2$.

@@

Since the phase rotation method operates on the same data as before, there will be no reduction in the acquisition frame rate by applying this method. In addition, since the same number of covariance matrices are constructed and inverted, there is no extra cost involved here as well. The only additional cost will be the multiplication of the extra steering vectors with the inverted covariance matrix. If the equation $\mat{\hat{R}}\vec{a} = \vec{b}$, as in \cite{Asen}, is solved using the Gauss-Jordan algorithm, we can solve the problem for all steering vectors at once by row reducing the matrix $[\mat{\hat{R}},  \vec{a}_{\theta_1}, \vec{a}_{\theta_2}, \cdots \vec{a}_{\theta_{N_\theta}}]$ to the following form $[\mat{I},  \vec{b}_{\theta_1}, \vec{b}_{\theta_2}, \cdots \vec{b}_{\theta_{N_\theta}}]$. This is still an $O(L^3)$ operation.

@@

In Fig\,\ref{fig:image_das} one point scatterer has been selected (white box), and then used to generate LSV-plots. \comment{Todo: Add lateral coherent interpolation as done for the simulations? As for now speckle is not fully resolved without it.} Figure \ref{fig:phantom_das} shows the resulting LSV-plot when using the DAS beamformer. The scalloping loss is less than 1 dB, and the system is therefore locally laterally shift-invariant. As a consequence point scatterers have a consistent appearance when subjected to motion \multimedia{Video DAS}.

Figure \ref{fig:phantom_capon_LSV} shows the same for the Capon beamformer without oversampling. In the frame where the point scatterer is at -6 degrees the receive beam direction is very close to the signal propagation vector. Hence, we get a high amplitude value at this location. In almost all the other frames the signal loss is massive (10 dB). The result is blinking point scatterers when subjected to motion \multimedia{Video Capon}.

Figure \ref{fig:phantom_capon_LSV_10PR} presents the result of applying ten times oversampling by phase rotation. We observe how the amplitude of the point scatterer is better preserved in all frames, and the scalloping loss is not far from our 1 dB threshold. Note that the oscillation along the diagonal does not indicate shift-variance. It is a result of lack of control when the probe was manually dragged across the phantom. In order to get a diagonal without peaks the probe has to be dragged across the phantom with a speed which matches the acquisition frame rate. In addition, due to the increase in lateral resolution the Capon beamformer requires smaller frame-to-frame movements than DAS. Looking at the video of Capon beamforming with ten times oversampling by phase rotation \multimedia{Capon + 10PR} there is no visible blinking anymore. We therefore conclude that the system is now locally laterally shift invariant.

@@

\section{Discussion}\label{sec:dis}
In this paper we have investigated the lateral shift-invariance property of the Capon beamformer when imaging moving objects. In Section \ref{sec:capon_LLSI} it was shown that a Capon beamformer with high resolution parameters is not locally laterally shift-invariant when subjected to motion. The root cause was found to be a too low lateral sampling density. In Section \ref{sec:methods} we therefore presented methods for improved shift-invariance based on oversampling. The most promising method, with respect to improved shift-invariance, reduction in acquisition frame rate and processing complexity, was based on phase rotation. In Section \ref{sec:res} we presented results showing that this method also works on \textit{in-vitro} data when the probe is subjected to motion.

%Note that in this paper we have investigated a single array configuration and a limited parameter range for the Capon beamformer. The goal was not to make a reference for which oversampling factor to use in different scenarios, but rather do a detailed investigation of the scalloping loss effect. We hope with this investigation that future readers will be able to find out them self what oversampling factor to use for their system.

If faced with an imaging scene containing bright point scatterers or sharp edges, one could wonder how large the oversampling factor needs to be in order to avoid visible scalloping losses. In his seminal paper, Cox \cite{Cox1973} derives formulas for the maximal achievable resolution for a given output SNR. For instance 47 dB is found to give a ten time improvement in resolution. We would therefore recommend to first use the dynamic range in the image as guidance. The amount of diagonal loading will also dictate the maximal achievable resolution since $10\log{d}$ can be interpreted as the input SNR seen by the Capon beamformer. Cox measures resolution as the 3 dB saddle point between two impinging monochromatic plane waves. For ultrasound data only a 50 \% increase in resolution was found in \cite[Fig.~7]{Asen} for a 30 dB output SNR when measured in a similar way as Cox (6 dB saddle point). However, sixteen times oversampling was required to make the imaging system shift invariant. In this paper more than ten times oversampling was found to be required for a 40 dB output SNR (after array gain) in simulations, however for the \textit{in-vitro} data ten times oversampling was enough for the system to be locally laterally shift-invariant. This might be a result of improved initial sampling (as discussed in Section \ref{sec:res}) and a higher initial noise level before diagonal loading was applied.

Relating the presented phase rotation method to earlier work we find that it is similar to a steering vector uncertainty set. However, where these set-based methods typically has a criterion for selecting the best output after applying all the steering vectors contained in the set, oversampling with phase rotation outputs all. It can also be thought of as a narrow band approach to parallel receive beamforming where Capon is applied on all receive beams.
